### 2015-01-21 paper.tex
cleanup

\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{extramarks}
\usepackage{pagecounting}
\usepackage{color}
\usepackage[round]{natbib}
\usepackage{lipsum}
\usepackage{enumerate}
%\usepackage{tipa}
%\usepackage{gb4e}
\usepackage{graphicx}
\usepackage{amsmath, amsthm}
%\usepackage{colortbl}
\usepackage{caption}
\usepackage[a4paper]{geometry}
\usepackage{courier}
\usepackage{listings}
\lstset{
         basicstyle=\footnotesize\ttfamily, 
         numberstyle=\tiny,          
         numbersep=5pt,             
         tabsize=2,                
         extendedchars=true,      
         breaklines=true,        
         showspaces=false,      
         showtabs=false,       
         xleftmargin=17pt,
         framexleftmargin=17pt,
         framexrightmargin=5pt,
         framexbottommargin=4pt,
         showstringspaces=false 
 }
 \lstloadlanguages{
         Haskell
 }

@@

Unit propagation magic!

@@

>>>>

\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{extramarks}
\usepackage{pagecounting}
\usepackage{color}
\usepackage[round]{natbib}
\usepackage{lipsum}
\usepackage{enumerate}
%\usepackage{tipa}
%\usepackage{gb4e}
\usepackage{graphicx}
\usepackage{amsmath, amsthm}
%\usepackage{colortbl}
\usepackage{caption}
\usepackage[T1]{fontenc}
\usepackage[a4paper]{geometry}
\usepackage{courier}
\usepackage{listings}
\lstset{
         basicstyle=\footnotesize\ttfamily, 
         numberstyle=\tiny,          
         numbersep=5pt,             
         tabsize=2,                
         extendedchars=true,      
         breaklines=true,        
         showspaces=false,      
         showtabs=false,       
         xleftmargin=17pt,
         framexleftmargin=17pt,
         framexrightmargin=5pt,
         framexbottommargin=4pt,
         showstringspaces=false 
 }
 \lstloadlanguages{
         Haskell
 }

@@

SAT solving is based on a technique called unit propagation:
a set of complex Boolean clauses is simplified, starting from unit
clauses, which consist of just a single variable, and working up to a solution, where all variables in the set
have a True or False value.

Let us show this with a small example.
We have a set of variables describing properties of people: \texttt{is\_under\_18},
\texttt{has\_driving\_licence} and \texttt{is\_drunk}.

% For instance, the following constraints describe rules for driving a car.
% \begin{verbatim}
% can_drive
% has_driving_licence
% is_under_18
% is_drunk
% \end{verbatim}

% Next, we represent a possible situation in the world. In order to be allowed to drive, a person must have a driving licence and they must not be drunk. This is represented by the clause $can\_drive ∧\wedge has\_driving\_licence ∧\wedge \neg is\_drunk$. Furthermore, from the world knowledge, we know that in order to have a driving licence, one must be over 18. This is represented by the second clause.

Clause 1 describes a possible situation in a world: someone is behaving in away that makes you think of a hypothesis ``either they are underage or
drunk''. This is represented as \texttt{is\_under\_18} $\vee$
\texttt{is\_drunk}. Just by this clause we can't solve the reason for the
behaviour: it can be one of them or both.

Clause 2 comes from the world knowledge that in order to have a
driving licence, one must be over 18. The form of the clause is
obtained by applying negation distributively over the more intuitive
clause $\neg$ (\texttt{has\_driving\_licence} $\wedge$
\texttt{is\_under\_18}): it is not possible to both have driving
licence and be under 18.

\begin{enumerate}
\item \texttt{is\_under\_18} $\vee$ \texttt{is\_drunk}
\item ($\neg$ \texttt{has\_driving\_licence}) $\vee$ ($\neg$ \texttt{is\_under\_18})
\item \texttt{has\_driving\_licence}
\end{enumerate}

Clause 3 provides us some undeniable evidence that
the person has a driving licence. By this clause, 
we know that the variable \texttt{is\_under\_18} is not true.
Knowing that, we can eliminate \texttt{is\_under\_18} from the first
clause. 
The solution is below: the three variables must have these values in order to
satisfy the constraints.

\begin{itemize}
\item \texttt{is\_drunk} = True
\item \texttt{is\_under\_18} = False
\item \texttt{has\_driving\_licence} = True
\end{itemize}

SAT solving is being used for many applications where

@@

\subsection{Complex conditions}
Conditions constructed by OR are represented as a sequence, and AND as
parallel. In Haskell, I simply use nested lists: everything inside the
outer list is subject to OR, and within the lists AND. This is best
demonstrated by an example:

\begin{itemize}
\item \texttt{OR (OR C1 C2) C3} becomes \texttt{[[C1], [C2], [C3]]} 
\item \texttt{AND (AND C1 C2) C3} becomes \texttt{[[C1, C2, C3]]} 
\item \texttt{AND (OR C1 C2) C3} becomes \texttt{[[C1,C3], [C2,C3]]}
\end{itemize}

\subsection{Conflicts}
If you try to use conflicting rules, you get user error: mzero.

\subsection{Example run}

We show an example run for the sentence \emph{They are both happy and go to the house}. This is
taken from a real output of a morphological analyser, and was
initially ambiguous in the following ways:

\begin{itemize}
\item[] are
\begin{itemize}
\item[] \texttt{<''be'', V, Pres>}
\item[] \texttt{<''are'', N, Sg>}
\end{itemize}

\item[] both
\begin{itemize}
\item[] \texttt{<''both'', CoordConj>}
\item[] \texttt{<''both'',Det>}
\item[] \texttt{<''both'',Pron>}
\end{itemize}

\item[] go
\begin{itemize}
\item[] \texttt{<''go'', V, Inf>}
\item[] \texttt{<''go'', V, Pres>}
\item[] \texttt{<''go'', N, Sg>}
\end{itemize}

\item[] to
\begin{itemize}
\item[] \texttt{<''to'', Prep>}
\item[] \texttt{<''to'', Adv>}
\end{itemize}

\item[] house
\begin{itemize}
\item[] \texttt{<''house'', V, Inf>}
\item[] \texttt{<''house'', V, Pres>}
\item[] \texttt{<''house'', N, Sg>}
\end{itemize}
\end{itemize}

\begin{verbatim}
Tag sequence:
1: [Lem "they",Pron,Subj,Nom,P3]
2: [Lem "are",N,Sg]
3: [Lem "both",Pron]
4: [Lem "happy",Adj]
5: [Lem "and",CoordConj]
6: [Lem "go",V,Inf]
6: [Lem "go",V,Pres]
7: [Lem "to",Prep]
8: [Lem "the",Det]
9: [Lem "house",N,Sg]
\end{verbatim}
